---
title: "Interactive Report: Airway Data"
date: March 5, 2026
bibliography: https://gist.githubusercontent.com/johnsonra/121eb61bf4b09e4258b78e341b1819e9/raw/51d87906d0b5824f9809f667008046665b815649/references.bib
server: shiny
---

## Study Summary

The following is a summary of the RNA-Seq analysis of airway smooth muscle cells as reported by @himes_rna-seq_2014.

### Background & Objective

Asthma is a chronic inflammatory disease of the airways, often managed using glucocorticoids (inhaled steroids). While these drugs are effective at reducing inflammation, the specific molecular mechanisms by which they act on Airway Smooth Muscle (ASM) cells—the cells responsible for airway constriction—were not fully mapped prior to this study.

The researchers aimed to use RNA-seq to:

1.  Characterize the global transcriptional response of ASM cells to glucocorticoids.

2.  Identify novel genes that might mediate the anti-inflammatory effects of these drugs.

### Experimental Design

The study utilized four primary human ASM cell lines. These lines were treated with:

-   Dexamethasone: A potent synthetic glucocorticoid (1 micromolar for 18 hours).

-   Vehicle Control: To establish a baseline of gene expression.

The researchers performed high-throughput RNA sequencing (RNA-seq), generating approximately 30 million reads per sample. This allowed for a high-resolution view of the transcriptome compared to older microarray technologies.

### Key Results

The analysis revealed a robust and consistent transcriptional shift across all four cell lines:

-   Differential Expression: See the volcano below for a visualization of differentially expressed genes.

```{r}
#| echo: false
sliderInput("thresh", "Significance threshold (adjusted p-value):",
            min = 0.0001, max = 0.1, value = 0.05, step = 0.0001)
```

```{r}
#| echo: false
plotOutput("volcanoPlot")
```

```{r}
#| context: server
library(ggplot2)
library(dplyr)
library(plotly)

top2000 <- read.csv("https://gist.githubusercontent.com/johnsonra/121eb61bf4b09e4258b78e341b1819e9/raw/9d4d1650713378ce59e06a21b62f940433100a81/top_2000.csv")

output$volcanoPlot <- renderPlot({
  thresh <- input$thresh

  top2000 |>
    mutate(nl10p = -log10(padj),
           sig = padj < thresh) |>

    ggplot(aes(log2FoldChange, nl10p, color = sig)) +
    geom_point(alpha = 0.3) +

    scale_y_sqrt() +
    scale_color_manual(values = c('black', 'orange')) +
    theme_minimal() +
    labs(x = bquote("log"[2] ~ "Fold Change"),
         y = bquote("-log"[10] ~ "p"),
         color = paste("Adjusted p <", thresh))
})
```

-   Identification of CRISPLD2: The most notable finding was the high induction of the gene CRISPLD2 (Cysteine-Rich Secretory Protein LCCL Domain Containing 2). This gene was previously known to be involved in lung development but hadn't been strongly linked to the glucocorticoid response in asthma.

-   Functional Enrichment: Down-regulated genes were heavily involved in inflammatory pathways, including those related to cytokine activity and Chemokine signaling (e.g. *CCL2*, *CXCL12*).

-   Validation: The researchers used siRNA "knockdown" experiments to show that when *CRISPLD2* is inhibited, ASM cells produce more pro-inflammatory cytokines, suggesting that *CRISPLD2* is a key "brake" on inflammation.

### References
